Staged Abstract Interpreters

### 2019-06-30 paper/10_intro.tex
misc

In this paper, we propose an abstraction-without-regret approach to
eliminate those performance penalties for abstract interpreters
written in purely functional programming style. In short, our approach
borrows ideas from program specialization and high-performance DSLs,
and applies them to a particular kind of meta-programs, abstract
interpreters.
1) Using multi-stage programming (the Lightweight Modular Staging
framework \cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
specialize the abstract interpreter with respect to an input program
and then generate efficient low-level code that does the actual
analysis. The specialization can also eliminate the effect layers in the generated code.
2) We use a tagless-final approach to embed the target
language interpreter in a high-level host language (Scala), allowing
the user to implement different semantics modularly, including the
abstract semantics and staging, without changing the main generic
interpreter. Together with the multi-stage programming and type-level
annotations, this design allows user to derive staged abstract
interpreters without intrusive changes to the unstaged one. Therefore,
our approach is no regret in the sense of both performance and least
engineering efforts to achieve such performance. We elaborate these
two main ideas in detail as follow.

>>>>

In this paper, we propose an abstraction-without-regret approach to
eliminate those performance penalties for abstract interpreters
written in purely functional programming style. In short, our approach
borrows ideas from program specialization and high-performance DSLs,
and applies them to a particular kind of meta-programs, abstract
interpreters.
1) Using multi-stage programming (the Lightweight Modular Staging
framework \cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
specialize the abstract interpreter with respect to an input program
and then generate efficient low-level code that does the actual
analysis. The specialization can also eliminate the effect layers in the generated code.
2) We use a tagless-final approach to embed the target
language interpreter in a high-level host language (Scala), allowing
the user to implement different semantics modularly, including the
abstract semantics and staging, without changing the main generic
interpreter. Together with the multi-stage programming and type-level
annotations, this design allows user to derive staged abstract
interpreters without intrusive changes to the unstaged one, thus the soundness is preserved. 
Therefore, our approach is no regret in the sense of both performance and least
engineering efforts to achieve such performance. We elaborate these
two main ideas in detail as follow.
